\input{../model}


\title{
    %\vspace{-2cm}
    \unilogo[0.3\textwidth]\\[0.9cm]
    S08: File System
}
\student{
    A. Schaller\\
    \texttt{16-896-375}
}
\prof{Dr. Luggen Michael}

\seriesnumber{8}
\author{\thestudent\theprof}
\lhead{\unilogo}
\rhead{\thestudent\\\theseriesnumber}

% \usepackage[backend=bibtex, style=numeric, citestyle=numeric, sorting=none, backref, backrefstyle=none]{biblatex}
% use "biber report" command to generate

% \addbibresource{biblio.bib}


% \usepgfplotslibrary{fillbetween}
\pgfplotsset{compat=newest}
% \usepackage{fontawesome}

\lstset{
    language=C,
    basicstyle=\footnotesize,
    stepnumber=3,
    firstnumber=1,
}

\lstset{literate=%
    {ß}{{\ss}}1
    {ç}{{\c{c}}}1
    {~}{{\textasciitilde}}1
}

\raggedbottom % To avoid stretching the items to fill a page
% ---------------------- DOCUMENT ----------------------------
\begin{document}

\maketitle


%
%
% ---------------------------------
\section{Create the function to\_upper(s)}

\question{Create the function to\_upper(s) which takes a String s as input and returns a String where all ASCII letters (a-z) are transformed to upper case, the other characters stay the same.}


\newquestion{Create one version which uses toupper(c) from \textless ctype.h\textgreater}

\begin{lstlisting}
#include <ctype.h>
#include <stdlib.h>
#include <string.h>

char * lib_upper (char * string) {
    int len = strlen(string);
    char * new_string = malloc(len);

    for (int i = 0; i < len; i++) {
        new_string[i] = toupper(string[i]);
    }

    return new_string;
}
\end{lstlisting}

\newquestion{Then extent the program with a function implementing your self toupper(c).
{\color{gray}\textbf{Hint}: To reimplement toupper(c) inspect closely the ASCII table of the lower, and upper case letters. The solution can be realized with simple arithmetic!}}

\begin{lstlisting}
char * own_upper (char * string) {
    int len = strlen(string);
    char * new_string = malloc(len);

    for (int i = 0; i < len; i++) {
        char c = string[i];
        new_string[i] = c;
        if (c >= 'a' && c <= 'z') {
            new_string[i] += 'A' - 'a';
        }
    }

    return new_string;
}
\end{lstlisting}

Small main function to print our results:
\begin{lstlisting}
#include <ctype.h>
#include <stdio.h>
#include <stdlib.h>
#include <string.h>

int main () {
    char s1[] = "Français";
    char s2[] = "spaß";
    printf("%s -> lib upper: %s\n", s1, lib_upper(s1));
    printf("%s -> lib upper: %s\n", s2, lib_upper(s2));

    printf("%s -> own upper: %s\n", s1, own_upper(s1));
    printf("%s -> own upper: %s\n", s2, own_upper(s2));

    return 1;
}
\end{lstlisting}


%
%
% ---------------------------------
\section{Write a program which writes to stdout and stderr.}

\newquestion{Write a tiny program which outputs alternating ``hey!'' to stdout and then ``ho!'' to stderr. In between the two output of the program shall sleep for 1 second.}

\begin{lstlisting}
#include <stdio.h>
#include <unistd.h>

int main () {
    while(1) {
        fprintf(stdout, "hey!\n");
        sleep(1);
        fprintf(stderr, "ho!\n");
        sleep(1);
    }
}
\end{lstlisting}


\newquestion{Write a bash command which writes all outputs of the stdout to the file hey.txt and all the output of the stderr to the file ho.txt}

\begin{lstlisting}[language={}]
./ex2 >> ./hey.txt 2>> ./ho.txt
\end{lstlisting}


%
%
% ---------------------------------
\section{Project P01: Linked Data In-Memory Store}

\question{After the project set-up with some dummy information try to define some data ``by hand'' and start to work on the match() function.}

\begin{itemize}
    \item \question{Define a data structure which you then fill with some example
    information statically (see the example from the \href{https://unifr.coursc.ch/\#projet}{project description}\footnotelink{https://unifr.coursc.ch/\#projet}{April 2021}). Make
    sure to define it outside the match() function.}

  \item \question{As a first step see if you can printf() all the triples which match the char* S variable.}
\end{itemize}

\end{document}